---
title: "Lab 4 Whitney Willoz"
author: "Whitney Willoz"
date: "10/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab 4
```{r}
library(readxl)
library(tidyverse)
defense <- read_excel('cyclonesFootball2020.xlsx', sheet='Defensive')
offense <- read_excel('cyclonesFootball2020.xlsx', sheet='Offensive')
bio <- read_excel('cyclonesFootball2020.xlsx', sheet='Biography')

str(defense)
str(offense)
str(bio)
```

# Part 1

1.)The player names and opponent names should be factors.
```{r}
defClean <- defense %>% 
  mutate(Name = factor(Name), Opponent_Opponent = factor(Opponent_Opponent))

offClean <- offense %>% 
  mutate(Name = factor(Name), Opponent_Opponent = factor(Opponent_Opponent))
  
bioClean <- bio %>% 
  mutate(Name = factor(Name))
```

2.) All offensive and defensive statistics should be numerical. The Weight column in biogrpahy should also be numeric. (Optional: Read ?across and learn to mutate multiple columns simultaneously using the across function in dplyr)
```{r}
defClean <- defClean %>% 
mutate(across(Tackles_Solo:Pass_PB, as.numeric))

offClean <- offClean %>% 
  separate(`Passing_CMP-ATT`, into = c("Passing_CMP","Passing_ATT")) %>%
mutate(across(Receiving_REC:Passing_INT, as.numeric)) 

bioClean <- bioClean %>% 
mutate(Weight = as.numeric(Weight))
```

3.)The Height column in biography should be numeric. (Hint: Use a unit that works well)
```{r}
bioClean <- bioClean %>%
  separate(Height, into = c("Heightf","HeightI"))%>%
  mutate(Heightf = as.numeric(Heightf), HeightI = as.numeric(HeightI))%>%
  mutate("Height (inches)" = (Heightf * 12)+ HeightI) %>%
  select(-Heightf,-HeightI)
```

4.)Add a column to both defense and offense indicating whether it is the first or the second game versus an opponent. We had a single game versus each opponent except for Oklahoma, which we played against twice. (Hint: You can use group_by(), mutate(), and row_number())
```{r}
defClean <- defClean %>% 
  group_by(Name)%>%
  mutate(First = !duplicated(Opponent_Opponent))

offClean <- offClean %>% 
  group_by(Name) %>%
  mutate(First = !duplicated(Opponent_Opponent))
```

Printed:
```{r}
str(defClean)
str(offClean)
str(bioClean)
```

# Part 2

1.)Reformat the offClean data frame into a tidy format using pivot_longer. The type of statistic (Receiving_REC for the number of receivings, Receiving_YDS for receiving yards, etc) is added as a new key column named stat.
```{r}
offClean <- offClean %>%
  pivot_longer(Receiving_REC:Passing_INT, names_to = "Stat_type",values_to = "Score")
```

2.)Obtain the player-level offensive statistics. Namely, for each player and each type of statistics, obtain the sum of the statistics in all games.
```{r}
playOffStats <- offClean %>%
  group_by(Name, Stat_type) %>% 
  summarize(Stat_sum = sum(Score))
playOffStats

```

3.)Show the distributions of the player-level offensive statistics. The distribution of each offensive statsitic should be shown in a separate facet. Explain the patterns you see.

Answer:I see that most of the statistics have a value for Brock Purdy and almost all players have a recieving_YDS value or a rushing_yds value. 

```{r}
ggplot(playOffStats, aes(x=Name,weight=Stat_sum, fill = Name)) + geom_bar() + facet_wrap(~Stat_type)+ coord_flip()
```

4.)Did ISU have better offense against Oregon or Oklahoma? Answer this question by creating a scatterplot with the x- and y-axes being the receiving yards of the ISU players against each of these two opponents. A large number of receiving yards indicates good offense.

Answer: It looks as though that we had a better offense against Oklahoma because we were able to get more Receiving yards on them. Using the first game comparison this is the results.
```{r}
#First time we played oklahoma
offClean %>%
 filter(Opponent_Opponent %in% c("Oregon", "Oklahoma"), First == TRUE, Stat_type == "Receiving_YDS")%>%
  group_by(Name,Opponent_Opponent)%>% 
  summarize(Stat_sum = sum(Score)) %>%
ggplot(aes(Opponent_Opponent, Stat_sum, color = Opponent_Opponent)) + geom_point()


```

Now turn our attention to the player biography.

4.)Separate the Hometown column into two columns, namely, the city and the state where the player comes from. Print the first few rows of these two columns. (Hint: look at the sep= argument of separate)

```{r}
bioClean <- bioClean %>%
  separate(Hometown, sep = "," , into = c("City","State"))

bioClean %>%
  select(City, State, Name) %>%
  head(5)

```

5.)How many players are there from each state?

```{r}
bioClean %>%
  group_by(State) %>%
  summarize(Players_per_State = n())
```

6.)How does the performance of quarterback Brock Purdy relate to the performance of the ISU team? Comment on both the defense and offense.

Answer:Overall it looks as if Brock Purdy's performance affects the whole team when it come down to these graphs. More so in Offense because if he isn't throwing the less receiving yards the rest of the team has and rushing for that matter. On defense, I see that if we are doing well on offense we want to also do well on defense.

```{r}
PurdyOffStats <- offClean %>%
  filter(Stat_type %in% c("Passing_YDS","Receiving_YDS","Rushing_YDS"))%>%
  group_by(Name, Opponent_Opponent, Stat_type) %>% 
  summarize(Stat_sum = sum(Score))

ggplot(PurdyOffStats, aes(x=Name,weight=Stat_sum, fill = Name)) + geom_bar() + facet_grid(Opponent_Opponent~Stat_type)+ coord_flip()

PurdyDefStats <- defClean %>%
  pivot_longer(Tackles_Solo:Pass_PB, names_to = "Stat_type", values_to = "Score")%>%
  group_by(Name, Opponent_Opponent, Stat_type) %>% 
  summarize(Stat_sum = sum(Score))
  
ggplot(PurdyDefStats, aes(x=Name,weight=Stat_sum, fill = Name)) + geom_bar() + facet_grid(Opponent_Opponent~Stat_type)+ coord_flip()

```

